ST563 Final Project

### 2021-04-20 DRAFT.Rmd
Added MLR code and descriptions.

```{r setup, include=FALSE}
# Set default chunk options
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, fig.align = "center",
                      out.height = '3.5in')

@@

# Regression Methods
## Multiple Linear Regression
NEED CODE FROM YUN
```{r mlr}
#### REGRESSION ####
lm.reg <- lm(quality ~ . - class, data = trn)

@@

Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables. The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables. Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors. The formula for a multiple linear regression model with $p$ predictor variables is as follows:

$$
\begin{aligned}
&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
& \quad \text{where for } i = n \text{ observations}, \\
& \quad y_i = \text{dependent variable (response)} \\
& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
& \quad \beta_0 = \text{y-intercept} \\
& \quad \beta_j = \text{slope coefficients for each variable} \\
& \quad \epsilon_i = \text{residuals (error term of model)} \\
\end{aligned}
$$

@@

  * Linear relationship between response and predictors
  * No collinearity between predictor variables
  * $y_i$'s are iid
  * $\epsilon_i \sim N(0, \sigma^2)$

To assess these assumptions, we looked at various diagnostics plots:
```{r lm.diag}
# Look at residual plot for nonlinearity
regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")

@@

# Stitch all the graphs together
grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
```

~ talk about residual plots here ~

@@

```{r reg.select, out.height='3.5in', eval = F}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
smre <- lm.sub %>% summary()

>>>>

```{r setup, include=FALSE}
# Set default chunk options
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = FALSE,
                      warning = FALSE, fig.align = "center",
                      out.height = '3.5in')

@@

# Regression Methods
## Multiple Linear Regression
Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables. The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables. Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors. The formula for a multiple linear regression model with $p$ predictor variables is as follows:

$$
\begin{aligned}
&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
& \quad \text{where observation } i = 1, ..., n \\
& \quad Y_i = \text{dependent variable (response)} \\
& \quad X_{ij} = \text{independent variables for } j = 1, ... , p \\
& \quad \beta_0 = \text{y-intercept} \\
& \quad \beta_j = \text{slope coefficients for each variable} \\
& \quad \epsilon_i = \text{residuals (error term of model)} \\
\end{aligned}
$$
Although it is possible to use every predictor in the MLR model, we instead use variable selection to reduce the number of predictors to six (`volatile.acidity`, `chlorides`, `pH`, `total.sulfur.dioxide`, `sulphates`, and `alcohol`) while maintaining predictive accuracy (see Variable Selection section for details). Applying the MLR with the reduced model, we get the following:

```{r mlr}
#### REGRESSION ####
lm.reg <- lm(quality ~ volatile.acidity + chlorides + pH + total.sulfur.dioxide + sulphates + alcohol, data = trn)
summary(lm.reg)

@@

The adjusted R-squared value 0.3633 suggests that these six predictors combined explain about 36.33% of the variance in `quality` of a wine.

@@

  * Linear relationship between response and predictors
  * No collinearity between predictor variables
  * $Y_i$'s are iid
  * $\epsilon_i \sim N(0, \sigma^2)$

To assess these assumptions, we looked at various diagnostics plots:
```{r lm.diag}
# Look at residual plot for nonlinearity
regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
                 aes(y = lm.reg$residuals, x = lm.reg$fitted)) + 
            geom_point(position = 'jitter') + labs(y = "Residuals", x = "Fitted Values")

@@

# Stitch all the graphs together
grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)

scale.res <- scale(lm.reg$residuals)
scale.res.2std <- mean((scale.res <= 1.96) * (scale.res >= -1.96))
```

Residuals indicate that our MLR assumptions are mostly satisfied. The residuals vs. fitted values plot (top left) has unusual stratified lines, but this is caused by the actual responses being integer values of 4, 5, 6, 7, etc. rather than continuous values. The plot does not show heteroscedasticity, and residuals are spread evenly around 0 over the entire range of fitted values. When residuals are standardized, as in the top right plot, we expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`. The Q-Q plot at bottom left compares the normality of residuals against a theoretical distribution. Except for a slight downward bias of the lowest 5% of residuals at 2 standard deviations below 0, the data fits a normal distribution almost perfectly. On the bottom right plot, high leverage points (those with leverage $\ge$ 0.03) are not dramatically biased, being spread both above and below the 0 residual. Cook's Distance top out at about 0.05, well below the generally accepted threshold of 1 needed to quality as highly influential. These plots taken together suggest that the data generally fulfills the requirements for modeling using MLR.

@@

```{r reg.select, out.height='3.5in', eval = T}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
smre <- lm.sub %>% summary()
